Expression QTL hotspot dissection with mediation analysis and pleiotropy testing

### 2019-08-02 analysis/paper/methods.Rmd
added results

$$
vec(Y) = X vec(B) + vec(G) + vec(E)
$${#eq:mvlmm}

For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. We performed a
two-dimensional, two-QTL scan over a genomic region that included the entirety of the
hotspot. We fitted bivariate linear mixed effects models (Equations [-@eq:mvlmm;-@eq:g;-@eq:e]) at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].

$$
G \sim MN_{n x 2}(0, K, V_g)
$${#eq:g}

$$
E \sim MN_{n x 2}(0, I_n, V_e)
$${#eq:e}

@@

$$
\hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
$$

$$
\hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
$$

$$
l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
$$

@@

$$
LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
$${#eq:lod}

$$
ll(\lambda_1, \lambda_2)
$$

@@

For each hotspot and each local trait - nonlocal trait pair, we also performed a
mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait. Each mediation analysis involved fitting four linear models (Equations~\ref{model1}, \ref{model2}, \ref{model3}, \ref{model4}) [@chick2016defining].

$$
Y = \alpha 1 + WC + E
$${#eq:model1}

$$
Y = XB + WC + E
$${#eq:model2}

$$
Y = \alpha 1 + WC + M\beta + E
$${#eq:model3}

$$
Y = XB + WC + M\beta + E
$${#eq:model4}

>>>>

\begin{equation}
vec(Y) = X vec(B) + vec(G) + vec(E)
(\#eq:mvlmm)
\end{equation}

For each hotspot, we tested all pairs of traits, regardless of whether the traits were local or nonlocal. We performed a
two-dimensional, two-QTL scan over a genomic region that included the entirety of the
hotspot. We fitted bivariate linear mixed effects models (Equations \@ref(), \@ref(), and \@ref()) at each ordered pair of markers in a two-dimensional grid [@boehm2019testing].

\begin{equation}
G \sim MN_{n x 2}(0, K, V_g)
(\#eq:g)
\end{equation}

\begin{equation}
E \sim MN_{n x 2}(0, I_n, V_e)
(\#eq:e)
\end{equation}

@@

\begin{equation}
\hat B = (X^T\hat \Sigma^{-1}X)^{-1}X^T\hat\Sigma^{-1}Y
(\#eq:Bhat)
\end{equation}

\begin{equation}
\hat \Sigma = \hat V_g \otimes \hat K + \hat V_e \otimes I_n
(\#eq:Sigmahat)
\end{equation}

\begin{equation}
l_{10}(\hat B, \hat \Sigma) = - \frac{1}{2}\log_{10}\left((Y - X\hat B)^T\hat \Sigma^{-1}(Y - X\hat B)\right) - n\log_{10}(2\pi) - \log_{10}|\hat \Sigma|
(\#eq:loglik)
\end{equation}

@@

\begin{equation}
LOD = \log_{10} \left(\frac{\max_{\lambda_1, \lambda_2}L(B, \Sigma, \lambda_1, \lambda_2)}{\max_{\lambda}L(B, \Sigma, \lambda, \lambda)} \right)
\end{equation}{#eq:lod}

\begin{equation}
ll(\lambda_1, \lambda_2)
\end{equation}

@@

For each hotspot and each local trait - nonlocal trait pair, we also performed a
mediation analysis to assess the extent which the local trait mediates the QTL effect on the nonlocal trait. Each mediation analysis involved fitting four linear models (Equations \@ref(eq:model1), \@ref(eq:model2), \@ref(eq:model3), \@ref(eq:model4)) [@chick2016defining].

\begin{equation}
Y = \alpha 1 + WC + E
(\#eq:model1)
\end{equation}

\begin{equation}
Y = XB + WC + E
(\#eq:model2)
\end{equation}

\begin{equation}
Y = \alpha 1 + WC + M\beta + E
(\#eq:model3)
\end{equation}

\begin{equation}
Y = XB + WC + M\beta + E
(\#eq:model4)
\end{equation}
